Cybersecurity in UK Higher Education?

### 2019-09-11 ITNow_TP.tex

Cybersecurity matters, breaches are legion and sadly are a common feature in the news. Universities are  addressing the demand for educated cybersecurity professionals.  There are calls for formal education to respond to this situation. This is reinforced by a wider focus on digital skills and computer science education reform, especially across the nations of the UK.

@@

\section*{Challenges} \label{sec}
% TC: pick a couple of these as examples?
\subsection*{Educational Resources}\label{sec:EDResource}
It is hard to determine what is actually delivered as part of a degree programme, but a reasonable proxy for this are recommended textbooks and other supporting resources.
\subsubsection*{The Case of SQL Injection}\label{sec:SQL}
SQL injection is a major weakness: number one in the Open Web Application Security Project (OWASP) Top 10 , and has been in the Top 10 since at least 2003. Database textbooks where analysed in 2019 ~\cite{Drop2019}.  There were seven such books, but three books accounted for the 36 of 44 universities. Five of the seven (30 of the 44) had no mention of SQL injection. On the other two, the more popular one has a seriously flawed discussion.

@@

The web abounds with informal resources, such as tutorials and code snippets. How good are these? One invesigation \cite{Unruhetal2017a}, took the top five search results from Google for six queries. Of these 30 tutorials, six had SQL injection weaknesses, and three had Cross-Site Scripting (Number 7 in OWASP's Top Ten) weaknesses. Searching for these fragments in PHP projects on GitHub found 820 instances of these fragments, of which 117 were verified manually to be vulnerable --- 80\% of which were vulnerable to SQL injection.

@@

\section*{What has the BCS been doing to promote the inclusion of cyber security in general computer science degrees? }

 The BCS has been using our accreditation process to mandate formal inclusion of information security (2010) and then cybersecurity (2013) within the programmes we accredit.

@@

Industry, higher education, government and the relevant professional bodies have collaborated on the production of a set of guidelines which are to the benefit of the various stakeholders and wider society as discussed in a previous issue of ITNow~\cite{Irons2016}. This work led to the production of reference guidelines (``{\emph{Cybersecurity Principles and Learning Outcomes}}'')  and established a baseline of common knowledge, example learning outcome domains for cybersecurity within the degree courses and guidance on embedding the concepts.

Since 2015, we have been expecting accredited degrees to be compliant with the cybersecurity guidelines. A full cycle of accreditation visits has not yet taken place following this inclusion in requirements. What is being observed is the majority of visited institutions have now either adjusted their curricula to extend the coverage of cybersecurity or have a plan in place to do so. However, a minority are requiring encouragement to do so.

From the start of the Autumn 2015 term, up to and including the Summer 2019 term, the BCS have carried out 82 accreditation visits (including five international visits (2 in South Africa and 1 in Brunei, Cyprus, and Ireland)) . The BCS identified action was required to address concerns related to cybersecurity at 23 institutions; thus, 59 institutions were already delivering cybersecurity in keeping with the BCS expectations.

Long-term actions (`{\emph{At Threshold}}' judgements) were expected from 14 institutions (six in 2015/16, three in 2016/17 and five in 2018/19). Thirteen of these judgments were across all programmes; one was specifically against a generalist masters programme only. This indicates that the BCS will expect adjustments to have taken place before the next accreditation visit. It was commonly the case that adjustments had been made to design of the programmes of study, however the adjusted programme had not yet been delivered so the evidence base was incomplete in terms of how cybersecurity was assessed.

Short term actions were required from nine institutions; the outcomes of these actions were as follows: ({\emph{i}}) of the 11 UG programmes involved all were approved `{\emph{At Threshold}}'; ({\emph{ii}}) of the nine UG programmes involved, eight were approved and one refused; ({\emph{iii}}) of the five UG programmes involved, all approved `{\emph{At Threshold}}'; and ({\emph{iv}}) of the 3 UG programmes involved, all 3 were refused; and ({\emph{v}}) a further 5 which at the time of writing the outcome is not known

@@

 \begin{quote}
``{\emph{The second year project provides an opportunity for exploring security aspects in depth with an industrial use case.}}''
\end{quote}
\begin{quote}
``{\emph{Hacktivity and related learning and teaching approaches}}''
\end{quote}
\begin{quote}
``{\emph{Cyber Security Centre which permeates both the course and supports external links and opportunities for students.}}''
\end{quote}

@@

The BCS are improving the cybersecurity capacity in the UK by enhancing the education experience of the graduates on BCS Accredited degree programmes. If you want to reassured the graduates you recruit have some cyber knowledge then seek a BCS accredited degree. There are challenges ahead in terms of enhancing the resources upon which the degree provision depends.

>>>>

Cybersecurity matters! Breaches are legion and sadly are a common feature in the news. Universities are  addressing the demand for educated cybersecurity professionals.  There are calls for formal education to respond to this situation. This is reinforced by a wider focus on digital skills and computer science education reform, especially across the nations of the UK.

@@

\section*{Challenges} \label{sec}
% TC: pick a couple of these as examples?
\subsection*{Educational Resources}\label{sec:EDResource}
It is hard to determine what is actually delivered as part of a degree programme, but a reasonable proxy for this are recommended textbooks and other supporting resources.
\subsubsection*{The Case of SQL Injection}\label{sec:SQL}
SQL injection is a major weakness: number one in the Open Web Application Security Project (OWASP) Top 10, and has been in the Top 10 since at least 2003. Database textbooks were analysed in 2019 ~\cite{Drop2019}.  There were seven such books, but three books accounted for the 36 of 44 universities. Five of the seven (30 of the 44) had no mention of SQL injection. On the other two, the more popular one has a seriously flawed discussion.

@@

The web abounds with informal resources, such as tutorials and code snippets. How good are these? One investigation \cite{Unruhetal2017a}, took the top five search results from Google for six queries. Of these 30 tutorials, six had SQL injection weaknesses, and three had Cross-Site Scripting (Number 7 in OWASP's Top Ten) weaknesses. Searching for these fragments in PHP projects on GitHub found 820 instances of these fragments, of which 117 were verified manually to be vulnerable --- 80\% of which were vulnerable to SQL injection.

@@

\section*{What has the BCS been doing to promote the inclusion of cybersecurity in general computer science degrees? }

 The BCS has been using the accreditation process to mandate formal inclusion of information security (2010) and then cybersecurity (2013) within the programmes we accredit.

@@

Industry, higher education, government and the relevant professional bodies have collaborated on the production of a set of guidelines which are to the benefit of the various stakeholders and wider society as discussed in a previous issue of ITNow~\cite{Irons2016}. This work led to the production of reference guidelines (``{\emph{Cybersecurity Principles and Learning Outcomes}}'')  and established a baseline of common knowledge, example learning outcome domains for cybersecurity within  degree programmes and guidance on embedding the concepts.

Since 2015, we have been expecting accredited degrees to be compliant with the cybersecurity guidelines. Universties are visited on a quinquennial basis and a full cycle of accreditation visits has not yet taken place following this inclusion in requirements. What is being observed is that the majority of visited institutions have now either adjusted their curricula to extend the coverage of cybersecurity or have a plan in place to do so. However, a minority are requiring encouragement to do so.

From the start of the Autumn 2015 term, up to and including the Summer 2019 term, the BCS has carried out 82 accreditation visits including five international visits (2 in South Africa and 1 in Brunei, Cyprus, and Ireland) . The BCS identified action was required to address concerns related to cybersecurity at 23 institutions; thus, 59 institutions were already delivering cybersecurity in line with the BCS expectations.

Long-term actions (`{\emph{At Threshold }}' judgements) were expected from 14 institutions (six in 2015/16, three in 2016/17 and five in 2018/19). 13 of these judgments were across all programmes; one was specifically against a generalist masters programme only. This indicates that the BCS will expect adjustments to have taken place before the next accreditation visit. It was commonly the case that adjustments had been made to design of the programmes of study, however, the adjusted programme had not yet been delivered so the evidence base was incomplete in terms of how cybersecurity was assessed.

Short term actions were required from nine institutions; the outcomes of these actions were as follows: ({\emph{i}}) of the eleven UG programmes involved all were approved `{\emph{At Threshold}}'; ({\emph{ii}}) of the nine UG programmes involved, eight were approved and one refused; ({\emph{iii}}) of the five UG programmes involved, all were approved `{\emph{At Threshold}}'; and ({\emph{iv}}) of the three UG programmes involved, all were refused; and ({\emph{v}}) a further five which at the time of writing the outcome is not known

@@

 \begin{quote}
``{\emph{The second-year project provides an opportunity for exploring security aspects in depth with an industrial use case.}}''
\end{quote}
\begin{quote}
``{\emph{Hacktivity and related learning and teaching approaches}}''
\end{quote}
\begin{quote}
``{\emph{Cyber Security Centre which permeates both the course and supports external links and opportunities for students.}}''
\end{quote}

@@

The BCS is improving the cybersecurity capacity in the UK by enhancing the education experience of the students on BCS accredited degree programmes. If you want to be reassured that the graduates you recruit have some cyber knowledge, then seek a BCS accredited degree. There are challenges ahead in terms of enhancing the resources upon which the degree provision depends.
